# Practicum 1

## How to Store Data

Data is commonly stored in spreadsheets.

* Columns are variables of interest
* Rows are observations.

Example: A dataset we'll call `iris` which has 150 observations of three species of iris. Each observation measured the length and width of both the petals and sepals.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
head(iris) %>% pander::pander()
```

Example: A dataset we'll call `GradeBook` that has records of how well a student performed on exams. I'll refer to this storage as the *wide* orientation.
```{r, message=FALSE, warning=FALSE, echo=FALSE}
readr::read_csv('./data-raw/GradeBook.csv') %>% head() %>% pander::pander()
```
Or I could have stored the information in the following manner, which I'll refer to as the *long* orientation.
```{r, message=FALSE, warning=FALSE, echo=FALSE}
readr::read_csv('./data-raw/GradeBook2.csv') %>% head() %>% pander::pander()
```


## Tableau

### Task 1: Dragging variables onto destination
First we'll make a scatterplot with some colors.

```{r, fig.height=5, echo=FALSE}
iris %>%
  ggplot( aes(y=Petal.Length, x=Sepal.Length, color=Species)) +
  geom_point() +
  labs(title = 'Sepal Length vs Petal Length by Iris Species',
       x= 'Sepal Length (cm)', y='Petal Length (cm)')
```

The iris data I used for this graph is available here:
data: https://raw.githubusercontent.com/dereksonderegger/141/master/data-raw/iris.csv


### Task 2: Modifying how a variable is displayed
https://raw.githubusercontent.com/dereksonderegger/141/master/data-raw/Lock5_GPAGender.csv

Modifying how a variable is displayed

* GPAGender
    * Make a Barchart with Gender and # Piercings
    * Turn # of Piercings into a discrete value.
    * Fix Gender Levels.  IFF( [GenderCode] = 1, 'Male', 'Female' )
    * Change the colors associated with each gender.

### Task 3: Reorder categorical variable levels

https://raw.githubusercontent.com/dereksonderegger/141/master/data-raw/warpbreaks.csv


```{r, fig.height=5, echo=FALSE}
warpbreaks %>%
  ggplot( aes( x=tension, y=breaks, color=wool) ) +
  geom_point() +
  facet_grid( . ~ wool) +
  labs(title='Number of breaks vs tension level (Low, Medium, High)')
```

