# Week 8

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```
```{r}
library(tidyverse)
```

## Project Goal

This class project is intended to have students experience the process of:
    
1. Asking questions that interest them.
2. Search for available data sets that are pertinent to the question.
3. Wrangle the data into a usable format.
4. Create graphic(s) that address the question of interest.
5. Using those graphics, ask follow-up questions and what data you would need to next get.


## Overview of Story

### Introduction
Because student projects will cover a wide variety of topics, we can't assume that the audience knows the necessary background information. Therefore, the introduction will introduce the audience to the topic and relevant concepts and terminology. 

The introduction section is all about the subject area of interest and NOT statistical or computational issues. This section should be written for an non-technical audience and avoid methodological details. 

### Methods
The purpose of this section is to give other researchers information about where the data was acquired, what transformations or modifications you made. Essentially this where we put the annoying details about how we did what we did. 

This section is often very dense and the main focus is the statistical and computational issues encountered and choices made to address those issues.

In the written presentation, you can be verbose enough that somebody could re-create the analysis you did. In the presentation, the amount of detail can be toned down.

### Results
In the results section, you present the graphs and analyses that have been created and major take-away points are clearly emphasized. In this section, only results that our data support should be brought up. 

The content of this is a mixture of the results of the analysis as well as how to interpret those results in the context of the subject area of interest. 

### Discussion 

In the discussion section, we focus on the implications of our analysis and what they mean in the context of our subject are of interest. Open-ended questions that the current data does not address can be brought up and the author/presenter should feel free to offer speculation. In this section, we are focused on the subject area and not the statistical/computational issues.

### References
This section documents any articles and/or database locations and access dates.


## Expectations
For this project, I expect that the presentation will be about 10 minutes and that the Introduction takes about 2 minutes, perhaps another 2 minutes for the methods, 5 minutes for the results, and a minute or 2 for the discussion and references.

For the paper aspect of the project, I expect projects to be a bit over 1000 words plus 3-5 appropriate graphs.


## Suggestions

