# Presentations 

```{r, echo=FALSE}
# Un-attach any packages that happen to already be loaded. In general this is unnecessary
# but is important for the creation of the book to not have package name-spaces
# fighting unexpectedly.
pkgs = names(sessionInfo()$otherPkgs)
if( length(pkgs > 0)){
  pkgs = paste('package:', pkgs, sep = "")
  for( i in 1:length(pkgs)){
    detach(pkgs[i], character.only = TRUE, force=TRUE)
  }
}
# Don't show the R code!
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```
```{r, warning=FALSE, message=FALSE}
# Chapter packages we will use
library(tidyverse)
```



## Introduction

Presenting your work to other people can take many different forms. 

1. Written documents are a communication medium that provides the opportunity for the most thorough discussion. The downside is that many people won't read the document if it is too overwhelming. Having a good document structure where people can jump around the document and read just what they are interested it can mitigate that. Because the author has no direct interaction with the reader, it is important to anticipate the questions that the readers might have and answer them. As a result, creating a written document takes quite a lot of time.

2. Oral presentations allow you to interact with the audience and that allows you to explain confusing points. The downside is that the audience can't skip around and re-visit points that they didn't understand. Most people won't interrupt to ask questions, so the presenter needs to read facial expressions to see how much confusion there is among the audience members.

3. Integration into other people's work. Well-created graphics that clearly explain a concept or idea will commonly be used by other folks. They should clearly attribute the graph to who ever created the graph, but it is useful to make sure that your graphics are in a state that they are stand-alone state that is interpretable. 
  
  
## Audience

Understanding the perspective of your audience is really important. I often think about presenting information to:

1. A high school educated individual without knowledge of the subject area I'm working with. Most popular news articles are written are written at this level and most of the details of the subject area are overlooked.

2. A college educated person with some knowledge of the subject. Articles and presentations at targeted at this level includes some large details and serve to keep individuals educated about fundamental ideas in a field.

3. An expert in the subject area. These articles and presentations are usually highly technical and very difficult to understand but are very important to the progression of the science.

As we write and create graphics, we want to be mindful of the different audiences and ideally a graph that was an quick introductory slide for a presentation given to an expert might be the final take-home message for the simplest audience level.

A really nice example of detail shifting is given by a YouTube series by Wired Magazine called [5 Levels](https://www.wired.com/video/series/5-levels). 

## Outline

1. Background information - Whatever background information your audience needs. When the audience is not highly educated about the topic, you need to include enough background to understand the context. A common mistake, however, is to include too much background and too much detail and overwhelm the audience. Try to be thoughtful giving just the details that absolutely necessary but it is probably better to give too much background information than too little. 

2. Data Source and Methods - To make a convincing discussion, it is necessary to describe the data that you are about to show and why it is appropriate to inform our discussion. It is also important to give enough information so that the reader/listener could access the data if it is publicly available. Any non-obvious transformations should be explained in sufficient detail as well. As your presentations include more sophisticated methods, some discussion of the analysis justification is appropriate here. The more sophisticated your audience is, the more important this section becomes.

3. Results - This section is the presents the data gathered and the graphs that support our insights. There might be many several graphs were we see the overall results and then follow it with  how those results vary by groups.

4. Discussion - This is the take home message that you want to leave the audience with. It is important to realize that the audience won't remember or understand everything written/said so making sure the take-home message is clear is important.

## Figures

1. Make sure all axis labels are understandable without explanation.
2. Make sure each figure has a title that describes what the reader what relationship is being explored. 
3. Highlight important feature that you want readers to see.

## Presentations

1. Projectors have much lower resolution and color contrast.
    a) Color palettes in need to be reduced because we'll only be able to distinguish between 4 to 5 separate colors. 
    b) Dark themes don't work well. Stick with dark text on a light background.
    
2. Slides 
    a) Minimalist Content Style- Just headers that serve as a speaker prompt about what to talk about. These give the audience a visual cue when you switch topics.
    b) Detailed Content Style - If you include details, it is important to keep them organized within the slide so that the audience isn't overwhelmed and not read anything. 
    c) No Wall of Text!
    d) It is fine to mix the content styles within a presentation.
    e) Regardless of the style, you should consider each slide should take approximately 1 to 2 minutes. Switching slides every 10 - 15 seconds just leaves the audience feel like you are rushing.
    f) Slides given in person don't have to use full sentences. However, if people will be reading the slides afterwards, then they should make sense when being read. In that case, the detail level has to go way up.


