# Week 10

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```
```{r, echo=FALSE}
library(tidyverse)
```

## Presentation Thoughts
In the last week, we had presentations from each student and they went quite well. However we want to have a discussion about the process that students underwent.

1. How much time was spent on data gathering? How much on data wrangling? How much on graphing?
    a) Weird file formats?
    b) Wide-vs-Long format?
    c) Can we systematically look for data? Domain knowledge is really important here.
    d) Hosting data is expensive, be thankful and think about the work somebody did.

2. Reproducible? : Given how your project stands right now, could future you (6 months from now) easily reproduce what you've done? Is there any single document that provides step-by-step instructions? Or are the steps scattered throughout some scripts and some Excel magic that you won't be able to figure out the order of operations?

2. For each student presentation, what graphic can you remember, now 2-5 days afterward? Is is one graph or multiple graphs that you remember?

3. How critical were annotations for the various graphics presented?

4. Scale axes: During any presentation, did you find yourself wondering what a scale axis was? Some presentations where consistent some where not.

5. Copying images from Tableau to PowerPoint:  Did this work well or did we get fuzziness?
    a) Vector graphic formats store shapes and paths of objects and can be re-sized without losing sharpness. Examples, .pdf and .eps. These are good for most graphs, but are not efficient for maps or images.
    b) Raster graphics make a pixel map of the image and cannot be re-sized without losing sharpness. Examples are .bmp, .tif, .gif, .jpg.  These are fine for maps or images, but should be avoided for graphs.


## Interactive Graphics
We've thought about interactive graphics as

1. A method to get viewers to interact and therefore pay attention.
2. A shiny whiz-bang doodad that is fun to fiddle with.

But it is more subtle than that.

Often interactive visualizations are all about changing the level of focus.

### Aggregation Level
1. Interaction is to choose the level of resolution about a certain question.
2. Typically much of the information is hidden until an aggregation level is selected. 
3. By zooming in and out of resolution, we keep the big picture in mind while allowing us to dig into as much detail as you want or need.
4. Examples
    a) [Google Maps](maps.google.com)
    b) [Google Maps Developer Documentation](https://developers.google.com/maps/documentation/maps-static/dev-guide)
    

### Highlight Level
1. Often we can show a high-level view of an entire topic. However there are too many details to be presented all at once and we want to suppress the details, but we want to see the relationship.
2. Examples
    a) [Bruce Springstein's History](http://duelingdatalarge.blogspot.com)
    b) [Game of Throne's Character Interactions](http://beta.wind-and-words.com)


### The "Me" Layer
Viewers tend to be very self-centric and a goal of most any story or analysis should be to explain how the topic affects the reader. By allowing a user to interact with and ask how does this affect:

1. Somebody like me.
2. My favorite team.
3. My state/town.
4. Examples:
    a) [NFL Elo Ratings](https://projects.fivethirtyeight.com/complete-history-of-the-nfl/)
    b) [The Birthday Problem](https://pudding.cool/2018/04/birthday-paradox/)


